Entrainment of low Mach number thermals in stratified domains

### 2019-05-01 tex/stratified_thermals.tex
some experiment section writing

\newcommand{\Div}[1]{\ensuremath{\nabla\cdot\left( #1\right)}}
\newcommand{\DivU}{\ensuremath{\nabla\cdot\bm{u}}}
\newcommand{\angles}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\KS}[1]{\ensuremath{\text{KS}(#1)}}
\newcommand{\KSstat}[1]{\ensuremath{\overline{\text{KS}(#1)}}}
\newcommand{\grad}{\ensuremath{\nabla}}
\newcommand{\RB}{Rayleigh-B\'{e}nard }
\newcommand{\stressT}{\ensuremath{\bm{\bar{\bar{\Pi}}}}}
\newcommand{\lilstressT}{\ensuremath{\bm{\bar{\bar{\sigma}}}}}
\newcommand{\nrho}{\ensuremath{n_{\rho}}}
\newcommand{\approptoinn}[2]{\mathrel{\vcenter{
	\offinterlineskip\halign{\hfil$##$\cr
	#1\propto\cr\noalign{\kern2pt}#1\sim\cr\noalign{\kern-2pt}}}}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% THEORY SECTION
\section{Theory}
\label{sec:theory}
The evolution of thermals as buoyant vortex rings has been well described in
the unstratified, Boussinesq limit for decades (as early as e.g., CITE, and
see \citet{lecoanet&jeevanjee2018} for other sources). Buoyant motions 
in the atmospheres of stars and planets are generally large enough to feel the
atmospheric stratification, and therefore a more thorough tratment of the
evolution of thermals in stratified domains is required to understand the nature
of thermal entrainment in nature.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% EXPERIMENT SECTION
\section{Experiment} 
\label{sec:experiment}
blah blah blah polytrope blah blah blah (do it with a general length scale).

blah blah blah 2D anelastic blah blah blah, nondimm'd for a direct comparison with
boussinesq.

blah blah blah 3D Fully compressible blah blah blah

blah blah blah thermal initial conditions, pressure equilibrium blah blah blah.

blah blah blah comparison of 2D and 3D blah blah blah. z v t, r v z, w v z. Also side-by-side
pictoral comparison with diff.

blah blah blah T(t) and derivative blah blah blah.

>>>>

\newcommand{\Div}[1]{\ensuremath{\nabla\cdot\left( #1\right)}}
\newcommand{\DivU}{\ensuremath{\nabla\cdot\bm{u}}}
\newcommand{\angles}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\td}[1]{\ensuremath{\widetilde{#1}}}
\newcommand{\KSstat}[1]{\ensuremath{\overline{\text{KS}(#1)}}}
\newcommand{\grad}{\ensuremath{\nabla}}
\newcommand{\RB}{Rayleigh-B\'{e}nard }
\newcommand{\stressT}{\ensuremath{\bm{\bar{\bar{\Pi}}}}}
\newcommand{\lilstressT}{\ensuremath{\bm{\bar{\bar{\sigma}}}}}
\newcommand{\nrho}{\ensuremath{n_{\rho}}}
\newcommand{\approptoinn}[2]{\mathrel{\vcenter{
	\offinterlineskip\halign{\hfil$##$\cr
	#1\propto\cr\noalign{\kern2pt}#1\sim\cr\noalign{\kern-2pt}}}}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% THEORY SECTION
\section{Theory}
\label{sec:theory}

\subsection{Phenomenological description of thermal evolution}
The initial conditions of the thermals studied in this work are spherical
entropy perturbations with zero initial velocity, as depicted in 
Fig.~\ref{fig:evolution_colormeshes}a. Buoyant forces cause these perturbations
to fall and develop into vortex rings, as depicted in 
Figs.~\ref{fig:evolution_colormeshes}c-f.

\begin{figure*}[t!]
    \includegraphics[width=\textwidth]{evolution_colormeshes.png}
    \caption{
    \label{fig:evolution_colormeshes} }
\end{figure*}

\subsection{Theoretical description of thermal evolution}
The evolution of thermals as buoyant vortex rings has been well described in
the unstratified, Boussinesq limit for decades (as early as e.g., CITE, and
see \citet{lecoanet&jeevanjee2018} for other sources). Buoyant motions 
in the atmospheres of stars and planets are generally large enough to feel the
atmospheric stratification, and therefore a more thorough tratment of the
evolution of thermals in stratified domains is required to understand the nature
of thermal entrainment in nature.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% EXPERIMENT SECTION
\section{Experiment} 
\label{sec:experiment}

\subsection{Polytropic Atmospheres}
We study an ideal gas whose equation of state is $P = \rho T$ and whose stratification
is a perfectly adiabatic polytrope,
\begin{gather}
T_0 = 1 + (\grad_{\text{ad}})(z - L_z) \\
\rho_0 = T_0^{m_{\text{ad}}},
\end{gather}
where $m_{\text{ad}} = (\gamma-1)^{-1}$, and where the adiabatic 
temperature gradient is $\grad_{\text{ad}} = - g / c_P$.

\subsection{Anelastic Equations}
The anelastic equations are \citep{lecoanet&all2014},
\begin{gather}
\td{\grad}\cdot\td{\bm{u}} = -\td{w}\partial_{\tilde{z}} \ln\rho_0 \\
\frac{D \td{\bm{u}}}{D \td{t}} = -\td{\grad} \td{\varpi} + g\frac{\td{S_1}}{c_P}\hat{z} + \frac{1}{\rho_0}\td{\grad}\cdot\left(\mu\td{\lilstressT}\right) \\
\frac{D \td{S_1}}{D\td{t}} = \frac{1}{\rho c_P}\td{\grad}\cdot\left(\kappa T_0 \td{\grad} \td{S_1}\right) + \frac{\mu}{\rho_0 T_0}\td{\sigma_{ij}}\partial_{\td{x_i}}\td{u_j}.
\end{gather}
where $D/D\td{t} = \partial/\partial \td{t} + \td{\bm{u}}\cdot\td{\grad}$ and the viscous stress tensor is
\begin{equation}
\td{\sigma_{ij}} = \left(\partial_{\td{x_i}}\td{u_j} + \partial_{\td{x_j}}\td{u_i} - \frac{2}{3}\delta_{ij}\td{\grad}\cdot\td{\bm{u}}\right).
\end{equation}

We nondimensionalize these equations as in \citet{lecoanet&jeevanjee2018} such that
the length scale is the diameter of the thermal and the velocity scale is the freefall
velocity,
\begin{equation}
\begin{split}
\td{\grad}\rightarrow(\td{L}_{th}^{-1})\grad, \qquad&
\td{S}_1 \rightarrow(\Delta\td{S})S_1,\\
\td{\bm{u}} \rightarrow (\td{u}_{th})\bm{u}, \qquad&
\td{\varpi} \rightarrow (\td{u}_{th}^2)\varpi,\\
\partial_{\tilde{t}} \rightarrow (\td{u}_{th}/\td{L}_{th})\partial_t,\qquad&
\end{split}
\end{equation}
with
\begin{equation}
\tilde{u}_{th}^2 = \frac{g \tilde{L}_{th} \Delta \tilde{s}}{c_P}, \qquad
\text{Re}_{\text{ff}} = \frac{\tilde{u}_{th} \tilde{L}_{th}}{\nu}, \qquad
\text{Pr}_{\text{ff}} = \frac{\tilde{u}_{th} \tilde{L}_{th}}{\chi}.
\end{equation}
The resulting equations are very similar to the boussinesq equations,
\begin{gather}
\DivU = -w \partial_z \ln\rho_0, \\
\begin{split}
\partial_t \bm{u} + \bm{u}\cdot\grad\bm{u} = \\
- \grad \varpi + S_1\hat{z} 
+ \frac{1}{\text{Re}_{\text{ff}}}\left[\grad^2 \bm{u} + \frac{1}{3}\grad(\DivU)\right] 
\end{split}\\
\begin{split}
\partial_t S_1 + \bm{u}\cdot\grad S_1 = \\
\frac{1}{\text{Re}_{\text{ff}}}\left(\frac{1}{\text{Pr}_{\text{ff}}\rho_0c_P }[\grad^2 S_1 + \partial_z\ln T_0 \cdot\partial_z S_1]\right.\\\left.
+ \frac{-(\grad_{ad})}{\rho_0 T_0}\sigma_{ij}\partial_{x_i}u_j \right)
\end{split}
\end{gather}

\subsection{Fully Compressible Equations}
The fully compressible equations, under the same assumptions as above, are
\begin{gather}
\frac{D \ln\rho}{Dt} + \DivU = 0 \\
\frac{D \bm{u}}{D t} = -\grad T - T\grad\ln\rho - g\hat{z} + \frac{1}{\rho}\grad\cdot\left(\mu\lilstressT\right) \\
\frac{D T}{Dt} + (\gamma-1)T\DivU  = \frac{1}{\rho c_V}\grad\cdot\left(\kappa \grad T\right) + \frac{\mu}{\rho c_V}\sigma_{ij}\partial_{x_i}u_j.
\end{gather}

\subsection{Initial Conditions}
We set the initial specific entropy perturbation as
\begin{equation}
S_1 = - A(1 - \text{erf}\left(\frac{r - r_{th}}{\delta}\right))/2,
\end{equation}
where in the FC cases $A = 10^{-4}$ and in the nondimensionalized AN cases,
$A = 1$. Here, $r = \sqrt{x^2 + y^2 + (z - z_0)^2}$, where $z_0 = Lz - 3*r_{th}$,
and $r_{th} = Lz/40$ is the radius of the thermal.

In the fully compressible equations, it is essential that the initial conditions be in pressure
equilibrium such that the low mach number nature of the thermal can be studied. In order to 
achieve this pressure balance, we set
\begin{equation}
\ln\rho_1 = S_1/c_P, \qquad T_1 = T_0(e^{-\ln\rho_1} - 1).
\end{equation}

\subsection{Solution for thermal evolution in a Polytrope}

\subsection{Verification of 2D Anelastic approximation}
blah blah blah comparison of 2D and 3D blah blah blah. z v t, r v z, w v z. Also side-by-side
pictoral comparison with diff.
